Self-assembly of colloid-cholesteric composites provides a possible route to switchable optical materials

### 2014-03-05 papers/2012-sandwich/response.txt

Dear Dr Huang,

Many thanks for sending us the referee reports on our manuscript, 
"Self-assembly of colloid-cholesteric composites: A route to switchable 
optical materials". We are glad that both referees found our work interesting.
We are also grateful for the suggestions for improvement, all of which
we have now implemented in the revised version. We hope the revised
manuscript can now be accepted for publication in Nature Communications.

@@

With best wishes,

K. Stratford, O. Henrich, J. S. Lintuvuori, M. E. Cates and D. Marenduzzo

As per the manuscript checklist, we include the captions for
Supplementary Movies here:

Supplementary Movie 1:
The final state of the simulation as shown in Fig.~1A. The movie
rotates the configuration to reveal more clearly its structure.

Supplementary Movie 2:
The final state of the simulation as shown in Fig.~1B. The movie
rotates the configuration to reveal more clearly its structure.

Supplementary Movie 3:
The final state of the simulation as shown in Fig.~1C. The movie
rotates the configuration to reveal more clearly its structure.

Supplementary Movie 4:
The final state of the simulation as shown in Fig.~1D. The movie
rotates the configuration to reveal more clearly its structure.

Supplementary Movie 5:
The configuration after the initial relaxation phase in Fig.~3. The movie
shows a rotation of the configuration. The distribution of the colloidal
particles is isotropic, with the particles forming clusters and having a
tendency to occupy positions at the intersection of the BPIII disclination
lines.

Supplementary Movie 6:
Configuration in a homogeneous electric field with strength
corresponding to ${\cal E}=0.8$. The movie
shows a rotation of the configuration.
A hexagonal pattern emerges where the particles form chains
situated at the vertices of the honeycomb-like structure. \\

@@

The paper provides insight into the effectiveness of including particles
into the blue phase as a means of controlling and adjusting the structure
and stability of the materials. The paper
is clearly written and the figures, paricularly the included 3D videos,
provide clear perspective on the modeled phases of these materials. The
results are of interest and useful to experimentalists working in the field. "

RESPONSE:
We are grateful to the reviewer for her/his overall positive view of our
work. We also found her/his comments constructive and useful to improve
our paper. We have addressed them as follows.

COMMENT:
"The paper would be even more useful if it included references to actual 
materials, both the liquid crystal host and the added particles and perhaps
even suggested a model system that would be a good starting point to study
experimentally. "

RESPONSE: 
We have fully taken this point on board.
In a new paragraph in the discussion section, we include six
new references specifically aimed at recent studies on experimental
nanoparticle liquid crystal composites. These are currently [34-39].
The paragraph also discusses the issue of a model system, and suggests
properties such a system may have to help match experiment and simulation.

### 2014-03-05 trunk/papers/2012-sandwich/CholColloids.tex
Final updates for response version
git-svn-id: https://ccpforge.cse.rl.ac.uk/svn/ludwig@2327 40876441-3496-0410-a613-a2fca35b2ae2

\title{Self-assembly of colloid-cholesteric composites provides
a route to switchable optical materials}

@@

\textcolor{red}{
The simulations presented here are representative of a liquid crystal
blue phase with unit cell size in the range of 100--500~nm \cite{mermin},
and particles with a diameter of around 50~nm. The simulations
address both the bulk structure, and the structure which forms when
the colloidal dispersion is confined in a narrow sandwich of width
comparable to the unit cell size~\cite{extrareference2}. It is seen
that a wide variety of structure is possible, and may be influenced
by parameters including the solid volume fraction, and the details
of the anchoring at the colloid surface (and the at the surface of the
confining walls). Further, evidence is found for metstable switching of
this structure in an applied electric field. Details of the
computational approach are set out in the Methods section, while
simulation parameters are reported in the Supplementary Information.
}

@@

\noindent
\textbf{Structure in bulk BPI.}
We show in Fig.~1 A-D the final state structures of four simulations in which we first equilibrated a stable BPI disclination network in bulk (with periodic boundary conditions) and then dispersed colloidal particles at random within it. Among dimensionless control parameters are the particle solid volume fraction $\phi$  ($\phi = 1\%$ or $5\%$); the ratio $r = R/\lambda$ of particle radius to the BP lattice parameter $\lambda$ (here $r\simeq 0.14$) and the ratio $w = WR/K$ where $W$ is the anchoring strength of the cholesteric at the particle surface, and $K$ is the elastic constant of the liquid crystal.
\textcolor{red}{
The anchoring strength is defined such that $W > 0$ imparts a preferred
orientation to the director field at the particle surface. This orientation
can be either normal to the surface or in a plane tangential to the surface,
a property related to surface chemistry in experimental systems.% but which
%can be imposed at will in simulations.
We choose} $w = 0.23, 2.3$ which are within typical experimental range~\cite{tiffany}. This parameter can be viewed as the ratio of the anchoring energy, $WR^2$, to the elastic energy scale for distortion $KR$.

@@

\medskip
\noindent
\textbf{Switching in electric field.}
An interesting question, relevant for applications to devices, is whether 
these structures can be modified, or switched, with electric fields.
Fig.~3 addresses this question, starting from amorphous Blue Phase III,
this time at a higher chirality. Dispersing colloidal particles in this networks
results in the particle occupying the available nodes, where the order
parameter is low (Fig.~3A; Supplementary Movie~5). Applying a high enough
field 
along the $x$ direction leads to a switching to a new phase, where colloidal 
arrays fit within a network with distorted hexagonal cells (Fig.~3B 
and Supplementary Fig.~7; Supplementary Movie~6). 
Remarkably, upon field removal this honeycomb structure does not find its 
way back to the 
configuration before switching on the field, but gets stuck into a metastable 
structure (Supplementary Fig.~7), with residual anisotropy along the field
switching 
direction. Cycling the field along the $x$ and $y$ directions leads to
distinct metastable structures, which are retained after field removal
(Supplementary Fig.~8 and Fig.~9).

@@

The effect of an electric field can be modelled by including the following
term in the bulk free energy density:
\begin{equation}
-\frac{\epsilon_a}{12\pi} E_{\alpha}Q_{\alpha\beta}E_{\beta},
\end{equation} 
where $E_{\alpha}$ are the components of the electric field, and 
$\epsilon_a$ (here assumed positive) is the dielectric anisotropy of the liquid
crystal. The strength of the electric field is quantified via
one further dimensionless number
\begin{equation}
{\cal E}^2 = \frac{27 \epsilon_a}{32 \pi A_0 \gamma} E_\alpha E_\alpha.
\end{equation}
The quantity ${\cal E}$ will be referred to as the reduced field strength.

@@

\bibitem{tanaka}
Araki, T., and Tanaka, H.,
Colloidal aggregation in a nematic liquid crystal: Topological arrest of
particles by a single-stroke disclination line,
{\it Phys. Rev. Lett.} {\bf 97}, 127801 (2006).

\bibitem{tiffany}
%T.~A. Wood, J.~S. Lintuvuori, A.~B. Schofield, D. Marenduzzo, W.~C.~K. Poon,
Wood, T.A., Lintuvuori, J.S., Schofield, A.B, Marenduzzo, D., and
Poon, W.C.K.,
A Self-Quenched Defect Glass in a Colloid-Nematic Liquid Crystal Composite,
{\it Science} {\bf 333}, 79--83 (2011).

@@

\bibitem{abbott}
%I.~H. Lin, D.~S. Miller, P.~J. Bertics, C.~J. Murphy, J.~J. de Pablo,
%N.~L. Abbott,
Lin, I.H., et al.,
Endotoxin-Induced Structural Transformations in Liquid Crystalline Droplet,
{\it Science} {\bf 332}, 1297--1300 (2011).

@@

\bibitem{mermin} Wright, D.C., and Mermin, N.D.,
Crystalline liquids: The blue phases, {\it Rev. Mod. Phys.} {\bf 61},
\textcolor{red}{385--432} (1989).

\bibitem{bp3} Henrich, O., Stratford, K., Cates, M.E., and Marenduzzo, D.,
 Structure of Blue Phase III of Cholesteric Liquid Crystals,
{\it Phys. Rev. Lett.} {\bf 106}, 107802 (2011).

@@

\bibitem{adriano}
Tiribocchi, A., Gonnella, G., Marenduzzo, D., Orlandini, E., and Salvadore, F.,
Bistable Defect Structures In Blue Phase Devices,
 {\it Phys. Rev. Lett.} {\bf 107}, 237803 (2011).

@@

%\bibitem{denniston2}
%C. Denniston, D. Marenduzzo, E. Orlandini, J.M. Yeomans,
%Lattice Boltzmann algorithm for three-dimensional liquid-crystal
%hydrodynamics,
%\textit{Philos. Trans. R. Soc. London, Ser. A} \textbf{362}, 1745--1754 (2004).

%\bibitem{codef} M. E. Cates, K. Stratford, R. Adhikari, P. Stansell,
%J.-C. Desplat, I. Pagonabarraga, A.J. Wagner,
%Simulating colloid hydrodynamics with lattice Boltzmann methods, 
%\textit{J. Phys. Condens. Matter} \textbf{16},  3903--3915 (2004).

@@

\centerline{\includegraphics[width=\textwidth]{text-fig1.jpg}}
\caption{\textbf{Bulk blue phase structures.} Snapshots of the states
obtained after dispersing
a suspension of colloidal nanoparticles within a cholesteric liquid
crystal in the BPI-forming region. (A-D) correspond to the case
of a dispersion of particles in a pre-equilibrated BPI phase;
(E-H) are obtained by dispersing colloids in an isotropic
phase and then quenching into the range where BPI is stable, leading to formation of an amorphous, BPIII-like, disclination network.
Structures correspond
to (A,E) $w=0.23$ and $\phi=1\%$;
(B,F) $w=0.23$ and $\phi=5\%$; 
(C,G) $w=2.3$ and $\phi=1\%$; and
(D,H) $w=2.3$ and $\phi=5\%$.
The anchoring of the director field to the colloidal surface is normal.
[For the full parameter list used to generate Figs.~1-3 see Supplementary
Information.] For clarity, only a portion of the simulation box is shown
(one eighth of the total volume); the full structures are shown in
Supplementary Figs.~2 and~3.}
\end{figure}

\begin{figure}
\centerline{\includegraphics[width=\textwidth]{text-fig2.jpg}}
\caption{\textbf{Confined blue phase structures.}
Snapshots of the steady states obtained when a dispersion of colloids in the isotropic phase is placed in a sandwich geometry, 
and then quenched into a regime where BPI is stable in the bulk. Both top and side views are provided.
The anchoring of the director field at the walls is normal for (A-D) and planar for (E-H). 
Structures correspond to (A,E) $w=0.23$ and $\phi=1\%$, 
(B,F) $w=0.23$ and $\phi=2\%$, 
(C,G) $w=2.3$ and $\phi=1\%$,
(D,H) $w=2.3$ and $\phi=2\%$.
As in Fig.~1, only one quarter of the simulation box is shown for clarity;
the full structures are shown in Supplementary Figs.~5 and~6.}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=0.5\textwidth]{col_disc_x_run1267.png}
\includegraphics[width=0.45\textwidth]{sq_x_run1267.png}\\
\includegraphics[width=0.5\textwidth]{col_disc_x_run1341.png}
\includegraphics[width=0.45\textwidth]{sq_x_run1341.png}
\end{center}
\caption{\textbf{Structure in electric field.}
Steady states obtained when a dispersion of colloids is inserted
into a thermodynamically stable bulk BPIII with and without applied
external electric field. Panel A shows the situation in zero field, and panel
C the situation when the applied field is in the $x$-direction
(perpendicular to the plane of the paper). Panels B and D show one
plane of the three-dimensional structure factor $S(\mathbf{k})$ computed
from the positions of the particles, with wavevector $k_x = 0$, corresponding
to A and C respectively. With field, the disclinations form a honeycomb
pattern with hexagonal ordering perpendicular to the field direction;
this is reflected in the structure factor.
Simulations use weakly anchoring particles $w=0.23$, and a volume
fraction of $\phi=5\%$ (see Supplementary Information for full list of
parameters). }
\end{figure}

>>>>

\title{Self-assembly of colloid-cholesteric composites provides
a possible route to switchable optical materials}

@@

\textcolor{red}{
The simulations presented here are representative of a liquid crystal
blue phase with unit cell size in the range of 100--500~nm \cite{mermin},
and particles with a diameter of around 50~nm. The simulations
address both the bulk structure, and the structure which forms when
the colloidal dispersion is confined in a narrow sandwich of width
comparable to the unit cell size~\cite{extrareference2}. It is seen
that a wide variety of structure is possible, and may be influenced
by parameters including the solid volume fraction, and the details
of the anchoring at the colloid surface (and the at the surface of the
confining walls). Further, evidence is found for metastable switching of
this structure in an applied electric field. Details of the
computational approach are set out in the Methods section, while
simulation parameters are reported in the Supplementary Information.
}

@@

\noindent
\textbf{Structure in bulk BPI.}
We show in Fig.~1 A-D the final state structures of four simulations in which we first equilibrated a stable BPI disclination network in bulk (with periodic boundary conditions) and then dispersed colloidal particles at random within it. Among dimensionless control parameters are the particle solid volume fraction $\phi$  ($\phi = 1\%$ or $5\%$); the ratio $r = R/\lambda$ of particle radius to the BP lattice parameter $\lambda$ (here $r\simeq 0.14$) and the ratio $w = WR/K$ where $W$ is the anchoring strength of the cholesteric at the particle surface, and $K$ is the elastic constant of the liquid crystal.
\textcolor{red}{
The anchoring strength is defined such that $W > 0$ imparts a preferred
orientation to the director field at the particle surface. This orientation
can be either normal to the surface or in a plane tangential to the surface,
a property related to surface chemistry in experimental systems.% but which
%can be imposed at will in simulations.
We choose} $w = 0.23, 2.3$ which are within typical experimental range~\cite{tiffany}. This parameter can be viewed as the ratio of the anchoring energy, $WR^2$, to the elastic energy scale for distortion~$KR$.

@@

\medskip
\noindent
\textbf{Switching in electric field.}
An interesting question, relevant for applications to devices, is whether 
these structures can be modified, or switched, with electric fields.
Fig.~3 addresses this question, starting from amorphous Blue Phase III,
this time at a higher chirality. Dispersing colloidal particles in this
network
results in the particle occupying the available nodes, where the order
parameter is low (Fig.~3A; Supplementary Movie~5). Applying a high enough
field 
along the $x$ direction leads to a switching to a new phase, where colloidal 
arrays fit within a network with distorted hexagonal cells (Fig.~3B 
and Supplementary Fig.~7; Supplementary Movie~6). 
Remarkably, upon field removal this honeycomb structure does not find its 
way back to the 
configuration before switching on the field, but gets stuck into a metastable 
structure (Supplementary Fig.~7), with residual anisotropy along the field
switching 
direction. Cycling the field along the $x$ and $y$ directions leads to
distinct metastable structures, which are retained after field removal
(Supplementary Fig.~8 and Fig.~9).

@@

The effect of an electric field can be modelled by including the following
term in the bulk free energy density:
\begin{equation}
-\frac{\epsilon_a}{12\pi} E_{\alpha}Q_{\alpha\beta}E_{\beta},
\end{equation} 
where $E_{\alpha}$ are the components of the electric field, and 
$\epsilon_a$ (here assumed positive) is the dielectric anisotropy of the liquid
crystal. The strength of the electric field is quantified via
one further dimensionless number
\begin{equation}
{\cal E}^2 = \frac{27 \epsilon_a}{32 \pi A_0 \gamma} E_\alpha E_\alpha.
\end{equation}
The quantity ${\cal E}$ is known as the reduced field strength.

@@

\bibitem{tanaka}
Araki, T., and Tanaka, H.,
Colloidal aggregation in a nematic liquid crystal: topological arrest of
particles by a single-stroke disclination line,
{\it Phys. Rev. Lett.} {\bf 97}, 127801 (2006).

\bibitem{tiffany}
%T.~A. Wood, J.~S. Lintuvuori, A.~B. Schofield, D. Marenduzzo, W.~C.~K. Poon,
Wood, T.A., Lintuvuori, J.S., Schofield, A.B, Marenduzzo, D., and
Poon, W.C.K.,
A self-quenched defect glass in a colloid-nematic liquid crystal composite,
{\it Science} {\bf 333}, 79--83 (2011).

@@

\bibitem{abbott}
%I.~H. Lin, D.~S. Miller, P.~J. Bertics, C.~J. Murphy, J.~J. de Pablo,
%N.~L. Abbott,
Lin, I.H., et al.,
Endotoxin-induced structural transformations in liquid crystalline droplet,
{\it Science} {\bf 332}, 1297--1300 (2011).

@@

\bibitem{mermin} Wright, D.C., and Mermin, N.D.,
Crystalline liquids: the blue phases, {\it Rev. Mod. Phys.} {\bf 61},
\textcolor{red}{385--432} (1989).

\bibitem{bp3} Henrich, O., Stratford, K., Cates, M.E., and Marenduzzo, D.,
 Structure of blue phase III of cholesteric liquid crystals,
{\it Phys. Rev. Lett.} {\bf 106}, 107802 (2011).

@@

\bibitem{adriano}
Tiribocchi, A., Gonnella, G., Marenduzzo, D., Orlandini, E., and Salvadore, F.,
Bistable defect structures In blue phase devices,
 {\it Phys. Rev. Lett.} {\bf 107}, 237803 (2011).

@@

@@

\centerline{\includegraphics[width=\textwidth]{text-fig1.jpg}}
\caption{\textbf{Bulk blue phase structures.} Snapshots of the states
obtained after dispersing
a suspension of colloidal nanoparticles within a cholesteric liquid
crystal in the BPI-forming region. (A-D) correspond to the case
of a dispersion of particles in a pre-equilibrated BPI phase;
(E-H) are obtained by dispersing colloids in an isotropic
phase and then quenching into the range where BPI is stable, leading to formation of an amorphous, BPIII-like, disclination network.
Structures correspond
to (A,E) $w=0.23$ and $\phi=1\%$;
(B,F) $w=0.23$ and $\phi=5\%$; 
(C,G) $w=2.3$ and $\phi=1\%$; and
(D,H) $w=2.3$ and $\phi=5\%$.
The anchoring of the director field to the colloidal surface is normal.
[For the full parameter list used to generate Figs.~1-3 see Supplementary
Information.] For clarity, only a portion of the simulation box is shown
\textcolor{red}{its linear extent being $\simeq 28R$}
(one eighth of the total volume);
the full structures are shown in
Supplementary Figs.~2 and~3.}
\end{figure}

\begin{figure}
\centerline{\includegraphics[width=\textwidth]{text-fig2.jpg}}
\caption{\textbf{Confined blue phase structures.}
Snapshots of the steady states obtained when a dispersion of colloids in the isotropic phase is placed in a sandwich geometry, 
and then quenched into a regime where BPI is stable in the bulk. Both top and side views are provided.
The anchoring of the director field at the walls is normal for (A-D) and planar for (E-H). 
Structures correspond to (A,E) $w=0.23$ and $\phi=1\%$, 
(B,F) $w=0.23$ and $\phi=2\%$, 
(C,G) $w=2.3$ and $\phi=1\%$,
(D,H) $w=2.3$ and $\phi=2\%$.
As in Fig.~1, only one quarter of the simulation box is shown for clarity
\textcolor{red}{(its horizontal extent being $\simeq 56R$ and vertical
extent $\simeq 24R$ in the narrow direction)};
the full structures are shown in Supplementary Figs.~5 and~6.}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=0.48\textwidth]{col_disc_x_run1267.png}
\includegraphics[width=0.43\textwidth]{sq_x_run1267.png}\\
\includegraphics[width=0.48\textwidth]{col_disc_x_run1341.png}
\includegraphics[width=0.43\textwidth]{sq_x_run1341.png}
\end{center}
\caption{\textbf{Structure in electric field.}
Steady states obtained when a dispersion of colloids is inserted
into a thermodynamically stable bulk BPIII with and without applied
external electric field. Panel A shows the situation in zero field, and panel
C the situation when the applied field is in the $x$-direction
(perpendicular to the plane of the paper). Panels B and D show one
plane of the three-dimensional structure factor $S(\mathbf{k})$ computed
from the positions of the particles, with wavevector $k_x = 0$, corresponding
to A and C respectively. With field, the disclinations form a honeycomb
pattern with hexagonal ordering perpendicular to the field direction;
this is reflected in the structure factor.
Simulations use weakly anchoring particles $w=0.23$, and a volume
fraction of $\phi=5\%$ (see Supplementary Information for full list of
parameters).
\textcolor{red}{The full simulation with a box size of $\simeq 56R$ is shown.}}
\end{figure}
